Optimizing Abstract$^{\text 2}$ Machines

### 2012-11-09 paper/paper.tex
Proof checkpoint.

\newtheorem{theorem}{Theorem}
% \usepackage{xltxtra}
% \setmonofont[Scale=MatchLowercase]{DejaVu Sans Mono}
%
\input{preamble}

@@

\begin{align*}
force(\clos{\mvar,\mexp,\menv},\msto) &= \{\clos{\mvar,\mexp,\menv}\}\\
force(\saddr\maddralt,\msto) &= \msto(\maddralt)\\
\ev{\svar{\mvar},\menv,\mkont,\msto} &\machstep\; 
\co{\mkont,\saddr{\menv(\mvar)},\msto} \\
\co{\kar[^\mcntr_\mlab]{\mexp,\menv,\maddr},v,\msto}
&\machstep\;
\ev[^\mcntr]{\mexp,\menv,\msto,\kfn[^\mcntr_\mlab]{\mval,\maddr}} \\
\mbox{where } &\mval' \in force(\mval,\msto) \\
bind^\mcntr_\mlab(\menv,\msto,\mvar,\mval) &= \menv[\mvar \mapsto
  \maddr], \msto\sqcup[\maddr \mapsto force(\mval, \msto)], \mcntr' \\
\mbox{where } \mcntr' &= \lfloor\mlab\mcntr\rfloor_k \\
              \maddr &= \mvar\mcntr'
\end{align*}

@@

Details of lazy machine given as a diff from the strict machine. \\
State space of $lazy-\widehat{CESK}^*_t$ varies just in
$\mathit{Value}$ (storable values do not change):
\begin{align*}
\mval \in \mathit{Value} &::= z \mid b \mid o \mid \\saddr{\maddr} \\
\end{align*}

@@

The property we show is that all states stay related through
reduction. The main difficulty is in showing the resulting stores are
in fact equal. We do this by showing each individual reduction takes
equal stores to equal stores, and that the union of all these
reductions is the full reduction.

\begin{theorem}
$\forall lcs,cs,lcs',\msto,\msto'$ if $\capprx{(lcs, \msto)}{(cs,\msto)}$
  and $(lcs,\msto) \machstep (lcs',\msto')$ then there exists $cs'$
  such that $(cs,\msto) \machstep (cs',\msto')$
\end{theorem}
\begin{proof}
...
\end{proof}

@@

%% Then we can show that if lazyconf ~ conf and lazyconf ==> lazyconf' then there exists a conf' such that conf ==> conf' and lazyconf' ~ conf'
%% Let lc in lazyconf.cs be arbitrary (call lazyconf.s, s)
%% By cases on lc:
%% <x r k d>:
%% Premises:
%% <x r k d> in conf.cs
%% Thus:
%% Since
%% <x r s k d> -->lazy   <addr(r(x)) s k d>
%% <x r s k d> -->strict <v s k d> where v in s(r(x))
%% By definition of force, v in force(addr(r(x)), s)
%% Thus the first rule of ~_s applies and the second rule of l~_s applies.

@@

%% <v ar^{l,lf,la}(e^le r a) d>:
%% Premises:
%% For all v' in force(v, s),
%%   <v' ar^{l,lf,la}(e^le r a) d> in conf.cs
%% Thus:
%% <v ar^{l,lf,la}(e^le r ka) d> -->lazy   <e^le r s\sqcup[lfd |-> force(v, s)] fn^{l,la}(lfd ka) d>
%% for all v' in force(v, s),
%% <v' ar^{l,lf,la}(e^le r ka) d> -->strict <e^le r s\sqcup[lfd |-> {v'}] fn^{l,la}(lfd ka) d>
%% Thus the union of all the strict stores will equal the lazy store.

%% <v fn^{l,la}(fa ka) d>:
%% Premises:
%% For all v' in force(v, s),
%%   <v' fn^{l,la}(fa ka) d> in conf.cs
%% Thus
%% <v fn^{l,la}(fa ka) d> -->lazy <e^fl r[x |-> xd'] s'\sqcup[xd' |-> s'(lad)] k d'>
%% where k in s(ka)
%%       clos(x e^fl r) in s(fa)
%%       d' = truncate(ld, K)
%%       s' = s\sqcup[lad |-> force(v, s)]
%% For all v' in force(v,s),
%% <v' fn^{l,la}(fa ka) d> -->strict <e^fl r[x |-> xd'] s'\sqcup[xd' |-> s'(lad)] k d'>
%% where k in s(ka)
%%       clos(x e^fl r) in s(fa)
%%       d' = truncate(ld, K)
%%       s' = s\sqcup[lad |-> {v'}]

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
% \usepackage{xltxtra}
% \setmonofont[Scale=MatchLowercase]{DejaVu Sans Mono}
%
\input{preamble}

@@

\begin{align*}
force(\msto,\clos{\mvar,\mexp,\menv}) &= \{\clos{\mvar,\mexp,\menv}\}\\
force(\msto,\saddr\maddralt) &= \msto(\maddralt)\\
\ev{\svar{\mvar},\menv,\mkont,\msto} &\machstep\; 
\co{\mkont,\saddr{\menv(\mvar)},\msto} \\
\co{\kar[^\mcntr_\mlab]{\mexp,\menv,\maddr},v,\msto}
&\machstep\;
\ev[^\mcntr]{\mexp,\menv,\msto,\kfn[^\mcntr_\mlab]{\mval,\maddr}} \\
\mbox{where } &\mval' \in force(\msto,\mval) \\
bind^\mcntr_\mlab(\menv,\msto,\mvar,\mval) &= \menv[\mvar \mapsto
  \maddr], \msto\sqcup[\maddr \mapsto force(\msto, \mval)], \mcntr' \\
\mbox{where } \mcntr' &= \lfloor\mlab\mcntr\rfloor_k \\
              \maddr &= \mvar\mcntr'
\end{align*}

@@

Details of lazy machine given as a diff from the strict machine. \\
State space of $lazy-\widehat{CESK}^*_t$ varies just in
$\mathit{Value}$ (storable values do not change):
\begin{align*}
\mval \in \mathit{Value} &::= z \mid b \mid o \mid \saddr{\maddr} \\
\end{align*}

@@

The property we show is that all states stay related through
reduction. The main difficulty is in showing the resulting stores are
in fact equal. We do this by showing each lazy reduction corresponds
to a set of strict reductions, for which the union of their output
stores equal the lazy store. Also for strict equality, every strict
reduction corresponds to a lazy reduction (immediate from the relation).

\begin{lemma}[Partition step]
$\forall lc, lc', cs, \msto, \msto'.$ if $\lapprx{lc}{\msto}{cs}$ and $wn(lc,\msto) \machstep wn(lc', \msto')$ then
$\exists cs'. \forall c \in cs. \exists c' \in cs', \msto^c. wn(c,\msto) \machstep wn(c', \msto^c)$ and $\msto' = \bigcup\limits_{c \in cs}{\msto^c}$ and $\lapprx{lc'}{\msto'}{cs'}$.
\end{lemma}
\begin{proof}
By cases on $wn(lc,\msto) \machstep wn(lc', \msto')$:
\begin{itemize}
\item{Case $\ev{\svar{\mvar}, \menv, \mkont} \machstep
            \co{\mkont, \maddr{\menv(\mvar)}, \msto}$: \\
      If $\mvar \in dom(\menv)$ and $\menv(x) \in dom(\msto)$, then $wn(lc, \msto)$ steps
      \begin{itemize} 
       \item{lazily to $\co{\saddr{\menv(\mvar)}, \msto, \mkont}$}
       \item{strictly to each of $\{\co{\mval, \msto, \mkont} \mid \mval \in \msto(\menv(\mvar))\}$}
      \end{itemize}
      By definition of ...}
\item{Case $\ev{\slit{\mlit}, \menv, \msto, \mkont} \machstep
            \co{\mkont, \mlit, \msto}$: \\
      Immeditate.}
\item{Case $\ev{\slam{\mvar}{\mexp}, \menv, \mkont} \machstep
            \co{\mkont, \clos{\mvar, \mexp, \menv}, \msto}$: \\
      Immediate}
\item{Case $\ev[^\mcntr]{\sapp[^\mlab]{\mexp_0}{\mexp_1}, \menv, \mkont} \machstep
            \ev[^\mcntr]{\mexp_0, \menv, \msto', \kar[^\mcntr_\mlab]{\mexp_1, \menv, \maddr}}$: \\
      Where $a, \msto' = \mathit{push}_\mlab^\mcntr(\msto,\mkont)$
      ...}
\item{Case $\ev[^\mcntr]{\sif[^\mlab]{\mexp_0}{\mexp_1}{\mexp_2}, \menv, \mkont} \machstep
            \ev[^\mcntr]{\mexp_0, \menv, \msto', \kif[^\mcntr]{\mexp_1, \mexp_2, \menv, \maddr}}$: \\
      Where $a, \msto' = \mathit{push}_\mlab^\mcntr(\msto,\mkont)$            
      ...}
\item{Case $\co{\kmt, \mval, \msto} \machstep \ans{\msto, \mval'}$: \\
      Where $\mval' \in force(\msto, \mval)$
      ...}
\item{Case $\co{\kar[^\mcntr_\mlab]{\mexp, \menv, \maddr}, \mval}$: \\
      ...}
\item{Case $\co{\kfn[^\mcntr_\mlab]{\mvalx{u}, \maddr}, \mval}$: \\
      ...}
\item{Case $\co{\kif[^\mcntr]{\mexp_0, \mexp_1, \menv, \maddr}, \strue}$: \\
      ...}
\item{Case $\co{\kif[^\mcntr]{\mexp_0, \mexp_1, \menv, \maddr}, \sfalse}$: \\
      ...}
\item{Case $\ap[^\mcntr_\mlab]{\clos{\mvar, \mexp, \menv}, \mval, \mkont}$: \\
      ...}
\item{Case $\ap[^\mcntr_\mlab]{\mop, \mval, \mkont}$: \\
      ...}
\end{itemize}
\end{proof}

\begin{theorem}[Laziness preserves precision]
$\forall lcs,cs,lcs',\msto,\msto'$ if $\capprx{(lcs, \msto)}{(cs,\msto)}$
  and $(lcs,\msto) \machstep (lcs',\msto')$ then there exists $cs'$
  such that $(cs,\msto) \machstep (cs',\msto')$
\end{theorem}
\begin{proof}
Let $lc \in lcs$ be arbitrary and let $\hat{c} \subseteq cs$ be the
smallest set such that $\lapprx{lc}{\msto}{\hat{c}}$. \\
...
\end{proof}

@@

%% Then we can show that if lazyconf ~ conf and lazyconf ==> lazyconf' then there exists a conf' such that conf ==> conf' and lazyconf' ~ conf'
%% Let lc in lazyconf.cs be arbitrary (call lazyconf.s, s)
%% By cases on lc:
%% <x r k d>:
%% Premises:
%% <x r k d> in conf.cs
%% Thus:
%% Since
%% <x r s k d> -->lazy   <addr(r(x)) s k d>
%% <x r s k d> -->strict <v s k d> where v in s(r(x))
%% By definition of force, v in force(s, addr(r(x)))
%% Thus the first rule of ~_s applies and the second rule of l~_s applies.

@@

%% <v ar^{l,lf,la}(e^le r a) d>:
%% Premises:
%% For all v' in force(s, v),
%%   <v' ar^{l,lf,la}(e^le r a) d> in conf.cs
%% Thus:
%% <v ar^{l,lf,la}(e^le r ka) d> -->lazy   <e^le r s\sqcup[lfd |-> force(s, v)] fn^{l,la}(lfd ka) d>
%% for all v' in force(s,v),
%% <v' ar^{l,lf,la}(e^le r ka) d> -->strict <e^le r s\sqcup[lfd |-> {v'}] fn^{l,la}(lfd ka) d>
%% Thus the union of all the strict stores will equal the lazy store.

%% <v fn^{l,la}(fa ka) d>:
%% Premises:
%% For all v' in force(s,v),
%%   <v' fn^{l,la}(fa ka) d> in conf.cs
%% Thus
%% <v fn^{l,la}(fa ka) d> -->lazy <e^fl r[x |-> xd'] s'\sqcup[xd' |-> s'(lad)] k d'>
%% where k in s(ka)
%%       clos(x e^fl r) in s(fa)
%%       d' = truncate(ld, K)
%%       s' = s\sqcup[lad |-> force(s,v)]
%% For all v' in force(s,v),
%% <v' fn^{l,la}(fa ka) d> -->strict <e^fl r[x |-> xd'] s'\sqcup[xd' |-> s'(lad)] k d'>
%% where k in s(ka)
%%       clos(x e^fl r) in s(fa)
%%       d' = truncate(ld, K)
%%       s' = s\sqcup[lad |-> {v'}]
